Summary of Dissertation Work Thus Far and Future Steps

### 2021-04-03 short-summary.Rmd
fixed a couple of typos

\begin{theorem}[Generalization to $K > 2$] 
\label{theorem2}
There exists a block diagonal matrix 
$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
$\{\lambda^{(kl)}\}_K$ and orthonormal matrix 
$U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
\end{theorem}

@@

$$\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$$

@@

In practice, however, we observe $A$ rather than $P$. Using the asymptotic
properties of the Adjacency Spectral Embedding (ASE) of the GRDPG
\cite{rubindelanchy2017statistical}, it can be shown that constructing 
$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}$
corresponding to between-community pairings going to 0 with probability 1.

@@

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
  \KwData{Adjacency matrix $A$, number of communities $K$, 
  hyperparameter $\lambda$}
  \KwResult{Community assignments $1, ..., K$}
    Find $V$, the matrix of eigenvectors of $A$ 
    corresponding to the $K (K + 1) / 2$ most positive 
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
      Assign $V_{-i} = \begin{bmatrix} 
      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
      Solve the LASSO problem 
      $c_i = \arg\min_{\beta} 
      \frac{1}{2} ||\sqrt{n} v_i - \sqrt{n} V_{-i} \beta||_2^2 + 
      \lambda ||\beta||_1$.\;
      Assign $\tilde{c}_i = \begin{bmatrix} 
      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
      \end{bmatrix}^\top$ such that the superscript is the index of 
      $\tilde{c}_i$.\;
    }
    Assign 
    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
    Compute the affinity matrix $B = |C| + |C^\top|$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$. 
\end{algorithm}

@@

So far, we've shown that there is a very convenient GRDPG representation of 
the PABM that makes community detection almost trivially simple. 
More specifically, we showed that an embedding of the PABM results in 
orthogonal subspaces and noise that goes to 0 with probability 1, which makes 
for a very straightforward application of SSC. It can also be shown that this 
type of embedding has the same properties of the SBM and DCBM, which are really
just special cases of the PABM.

Furthermore, all Bernoulli graphs can be
represented as a GRDPG, and it may be the case that other, more complicated 
graphs have nice ASE representations that lead to community detection methods. 
In particular, we are currently exploring the nature of RDPGs and GRDPGs that
result from latent positions for which each community lies on a manifold rather
than on a subspace. 
Much of this work is based on work by \citet{trosset2020learning} who showed 
convergence properties for graphs that result from latent positions that lie 
on a one-dimensional manifold.

>>>>

\begin{theorem}[Generalization to $K > 2$] 
\label{theorem2}
There exists a block diagonal matrix 
$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
$\{\lambda^{(kl)}\}_K$ and orthonormal matrix 
$U \in \mathbb{R}^{K^2 \times K^2}$ that is known 
for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
\end{theorem}

@@

$$\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k1)} & \cdots & \lambda^{(kK)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$$

@@

In practice, however, we observe $A$ rather than $P$. Using the asymptotic
properties of the Adjacency Spectral Embedding (ASE) of the GRDPG
\cite{rubindelanchy2017statistical}, it can be shown that constructing 
$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}_n$
corresponding to between-community pairings going to 0 with probability 1.

@@

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering for the PABM using LASSO.}
  \KwData{Adjacency matrix $A$, number of communities $K$, 
  hyperparameter $\lambda$}
  \KwResult{Community assignments $1, ..., K$}
    Find $V$, the matrix of eigenvectors of $A$ 
    corresponding to the $K (K + 1) / 2$ most positive 
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
      Assign $V_{-i} = \begin{bmatrix} 
      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
      Solve the LASSO problem 
      $c_i = \arg\min_{\beta} 
      \frac{1}{2} ||\sqrt{n} v_i - \sqrt{n} V_{-i} \beta||_2^2 + 
      \lambda ||\beta||_1$.\;
      Assign $\tilde{c}_i = \begin{bmatrix} 
      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
      \end{bmatrix}^\top$ such that the superscript is the index of 
      $\tilde{c}_i$.\;
    }
    Assign 
    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
    Compute the affinity matrix $B = |C| + |C^\top|$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$. 
\end{algorithm}

@@

So far, we've shown that there is a very convenient GRDPG representation of 
the PABM that makes community detection almost trivially simple. 
More specifically, we showed that an embedding of the PABM results in 
orthogonal subspaces and noise that converges to 0 with probability 1, which 
makes for a very straightforward application of SSC. It can also be shown that 
this type of embedding has the same properties for the SBM, DCBM, and Nested 
Block Model (NBM), which are really just special cases of the PABM
(see \citet{noroozi2021hierarchy}, in particular Fig 2.).

Furthermore, all Bernoulli graphs can be
represented as a GRDPG, and it may be the case that other, more complicated 
graphs have nice ASE representations that lead to community detection methods. 
In particular, we are currently exploring the nature of RDPGs and GRDPGs that
result from latent positions for which each community lies on a manifold rather
than on a subspace. 
Much of this work is based on work by \citet{trosset2020learning} who 
considered the RDPG with latent positions that lie on a manifold and developed 
methods for recovering the manifold from an adjacency matrix sampled from it.
